---
title: "ascotraceR workflow"
author: "Dr. Paul Melloy"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup}
library(ascotraceR)
library(stationaRy) # for weather data
library(data.table)
library(ggplot2)
library(nasapower)
library(foreach)
library(doParallel)
```


Download weather for turret

Potential trial to mimic SARDI fungicide trial in 2018.
[Description of the trial at GRDC publications](https://grdc.com.au/resources-and-publications/grdc-update-papers/tab-content/grdc-update-papers/2019/02/ascochyta-blight-in-intensive-cropping-of-pulses)

```{r get_power}
np_w <- get_power(community = "ag",
                  lonlat = c(138.832,-34.553),
                  pars = c("T2M","PRECTOTCORR","WD2M","WS2M"),
                  temporal_api = "hourly",
                  dates = c("2014-01-01","2014-12-30"),
                  time_standard = "UTC")
setDT(np_w)

```


```{r}
# Add station id
np_w[,id := "Turretfield"]

# set standard deviation of wd
np_w[,wdir_sd := 30]

f_dat <-
  format_weather(
    np_w,
    YYYY = "YEAR",
    MM = "MO",
    DD = "DY",
    hh = "HR",
    time_zone = "UTC",
    temp = "T2M",
    rain = "PRECTOTCORR",
    ws = "WS2M",
    wd = "WD2M",
    wd_sd = "wdir_sd",
    station = "id",
    lon = "LON",
    lat = "LAT"
  )
f_dat[,.(temp = mean(temp),
         rain = sum(rain)),
      by = .(YYYY,MM,DD)][MM >= 6 & MM <= 10,]
```

```{r}
T1 <- Sys.time()
ta1 <- trace_asco(weather = f_dat,
                  paddock_length = 10,
                  paddock_width = 20,
                  sowing_date = "2014-06-06",
                  harvest_date = "2014-09-17", # Guesstimated 
                  initial_infection = "2014-06-26",# estimated "Early in the season" 20 days after sowing
                  seeding_rate = 50, # Desi
                  primary_infection_foci = expand.grid(x = 1:20,
                                                       y = 1:10,
                                                       load = 5),
                  spores_per_gp_per_wet_hour = 0.22) # kyabra = 0.6
T2 <- Sys.time()
T2 - T1

s_ta1 <- summarise_trace(ta1)
td_ta1 <- tidy_trace(ta1)
```


```{r}
ggplot(data = subset(tidy_trace(ta1), i_day == 105),
       aes(x = x, y = y, fill = infectious_gp/susceptible_gp)) +
       geom_tile()



plot(y = s_ta1$infectious_gp/s_ta1$susceptible_gp,
     x = s_ta1$i_date,
     type = "l",
     ylab = "Proportion of infested growing points",
     xlab = "2014 Date")
```

How does increasing the amount of starting inoculum change the results?

```{r}
T1 <- Sys.time()
ta2 <- trace_asco(weather = f_dat,
                  paddock_length = 10,
                  paddock_width = 20,
                  sowing_date = "2014-06-06",
                  harvest_date = "2014-09-17", # Guesstimated 
                  initial_infection = "2014-06-26",# estimated "Early in the season" 20 days after sowing
                  seeding_rate = 50, # Desi
                  primary_infection_foci = expand.grid(x = 1:20,
                                                       y = 1:10,
                                                       load = 10),
                  spores_per_gp_per_wet_hour = 0.22) # kyabra = 0.6
T2 <- Sys.time()
T2 - T1

s_ta2 <- summarise_trace(ta2)
td_ta2 <- tidy_trace(ta2)
```


```{r}
ggplot(data = subset(td_ta2, i_day == 105),
       aes(x = x, y = y, fill = infectious_gp/susceptible_gp)) +
       geom_tile()



plot(y = s_ta2$infectious_gp/s_ta2$susceptible_gp,
     x = s_ta2$i_date,
     type = "l",
     ylab = "Proportion of infested growing points",
     xlab = "2014 Date")
```

How does changing the susceptibility change the results?

```{r}
T1 <- Sys.time()
ta3 <- trace_asco(weather = f_dat,
                  paddock_length = 10,
                  paddock_width = 20,
                  sowing_date = "2014-06-06",
                  harvest_date = "2014-10-30", # Guesstimated 
                  initial_infection = "2014-06-26",# estimated "Early in the season" 20 days after sowing
                  seeding_rate = 50, # Desi
                  primary_infection_foci = expand.grid(x = 1:20,
                                                       y = 1:10,
                                                       load = 5),
                  spores_per_gp_per_wet_hour = 0.6) # kyabra = 0.6
T2 <- Sys.time()
T2 - T1

s_ta3 <- summarise_trace(ta3)
td_ta3 <- tidy_trace(ta3)
```


```{r}
ggplot(data = subset(td_ta3, i_day == 105),
       aes(x = x, y = y, fill = infectious_gp/susceptible_gp)) +
       geom_tile()



plot(y = s_ta3$infectious_gp/s_ta3$susceptible_gp,
     x = s_ta3$i_date,
     type = "l",
     ylim = c(0,0.4),
     ylab = "Proportion of infested growing points",
     xlab = "2014 Date")
lines(y = s_ta2$infectious_gp/s_ta2$susceptible_gp,
     x = s_ta2$i_date,
     type = "l", col = "red")
```

What if we run this till christmas. Does the crop stop growing?

```{r}
T1 <- Sys.time()
ta4 <- trace_asco(weather = f_dat,
                  paddock_length = 10,
                  paddock_width = 20,
                  sowing_date = "2014-06-06",
                  harvest_date = "2014-12-30", # Guesstimated 
                  initial_infection = "2014-06-26",# estimated "Early in the season" 20 days after sowing
                  seeding_rate = 50, # Desi
                  primary_infection_foci = expand.grid(x = 1:20,
                                                       y = 1:10,
                                                       load = 5),
                  spores_per_gp_per_wet_hour = 0.2) # kyabra = 0.6
T2 <- Sys.time()
T2 - T1

s_ta4 <- summarise_trace(ta4)
td_ta4 <- tidy_trace(ta4)
```


```{r}
ggplot(data = subset(td_ta4, i_day == 105),
       aes(x = x, y = y, fill = infectious_gp/susceptible_gp)) +
       geom_tile()



plot(y = s_ta4$infectious_gp/s_ta4$susceptible_gp,
     x = s_ta4$i_date,
     type = "l",
     ylim = c(0,0.25),
     ylab = "Proportion of infested growing points",
     xlab = "2014 Date")

```


Test adding fungicide dates?

```{r}
T1 <- Sys.time()
ta5 <- trace_asco(weather = f_dat,
                  paddock_length = 10,
                  paddock_width = 20,
                  sowing_date = "2014-06-06",
                  harvest_date = "2014-10-30", # Guesstimated 
                  initial_infection = "2014-06-26",# estimated "Early in the season" 20 days after sowing
                  seeding_rate = 50, # Desi
                  primary_infection_foci = expand.grid(x = 1:20,
                                                       y = 1:10,
                                                       load = 5),
                  spores_per_gp_per_wet_hour = 0.6, # kyabra = 0.6
                  fungicide_dates = c("2014-07-07","2014-08-29","2014-09-07")) 
T2 <- Sys.time()
T2 - T1

s_ta5 <- summarise_trace(ta5)
td_ta5 <- tidy_trace(ta5)
```


```{r}
ggplot(data = subset(td_ta5, i_day == 105),
       aes(x = x, y = y, fill = infectious_gp/susceptible_gp)) +
       geom_tile()



plot(y = s_ta5$infectious_gp/s_ta5$susceptible_gp,
     x = s_ta5$i_date,
     type = "l",
     ylim = c(0,0.38),
     ylab = "Proportion of infested growing points",
     xlab = "2014 Date")
lines(y = s_ta3$infectious_gp/s_ta3$susceptible_gp,
     x = s_ta3$i_date,
     type = "l", col = "red")
abline(v = as.Date(c("2014-07-07","2014-08-29","2014-09-07")),col = "blue")

```


Test adding earlier fungicide dates?

```{r}
T1 <- Sys.time()
ta6 <- trace_asco(weather = f_dat,
                  paddock_length = 10,
                  paddock_width = 20,
                  sowing_date = "2014-06-06",
                  harvest_date = "2014-10-30", # Guesstimated 
                  initial_infection = "2014-06-26",# estimated "Early in the season" 20 days after sowing
                  seeding_rate = 50, # Desi
                  primary_infection_foci = expand.grid(x = 1:20,
                                                       y = 1:10,
                                                       load = 5),
                  spores_per_gp_per_wet_hour = 0.6, # kyabra = 0.6
                  fungicide_dates = c("2014-07-07","2014-07-22","2014-08-13")) 
T2 <- Sys.time()
T2 - T1

s_ta6 <- summarise_trace(ta6)
td_ta6 <- tidy_trace(ta6)
```


```{r}
ggplot(data = subset(td_ta6, i_day == 105),
       aes(x = x, y = y, fill = infectious_gp/susceptible_gp)) +
       geom_tile()



plot(y = s_ta6$infectious_gp/s_ta6$susceptible_gp,
     x = s_ta6$i_date,
     type = "l",
     ylim = c(0,0.38))
lines(y = s_ta3$infectious_gp/s_ta3$susceptible_gp,
     x = s_ta3$i_date,
     type = "l", col = "red")
lines(y = s_ta5$infectious_gp/s_ta5$susceptible_gp,
     x = s_ta5$i_date,
     type = "l", col = "maroon")
abline(v = as.Date(c("2014-07-07","2014-07-22","2014-08-13")),col = "blue")

```

Remove very early spray

```{r}
T1 <- Sys.time()
ta7 <- trace_asco(weather = f_dat,
                  paddock_length = 10,
                  paddock_width = 20,
                  sowing_date = "2014-06-06",
                  harvest_date = "2014-10-30", # Guesstimated 
                  initial_infection = "2014-06-26",# estimated "Early in the season" 20 days after sowing
                  seeding_rate = 50, # Desi
                  primary_infection_foci = expand.grid(x = 1:20,
                                                       y = 1:10,
                                                       load = 5),
                  spores_per_gp_per_wet_hour = 0.6, # kyabra = 0.6
                  fungicide_dates = c("2014-07-22","2014-08-13")) 
T2 <- Sys.time()
T2 - T1

s_ta7 <- summarise_trace(ta7)
td_ta7 <- tidy_trace(ta7)
```


```{r}
ggplot(data = subset(td_ta7, i_day == 105),
       aes(x = x, y = y, fill = infectious_gp/susceptible_gp)) +
       geom_tile()



plot(y = s_ta7$infectious_gp/s_ta7$susceptible_gp,
     x = s_ta7$i_date,
     type = "l",
     ylim = c(0,0.38))
lines(y = s_ta3$infectious_gp/s_ta3$susceptible_gp,
     x = s_ta3$i_date,
     type = "l", col = "red")
lines(y = s_ta6$infectious_gp/s_ta6$susceptible_gp,
     x = s_ta6$i_date,
     type = "l", col = "maroon")
abline(v = as.Date(c("2014-07-07","2014-07-22","2014-08-13")),col = "blue")

```


Permutate spores_per_wet_hour



```{r}
T1 <- Sys.time()
cl <- makeCluster(detectCores())
registerDoParallel(cl)

perm_sr <- foreach(sr = seq(from = 0.25,
                            to = 0.9,
                            by = 0.05),
                   w = f_dat,
                   .packages = "ascotraceR", 
                   .inorder=FALSE) %dopar% {
                     trace_asco(
                       weather = f_dat,
                       paddock_length = 10,
                       paddock_width = 20,
                       sowing_date = "2014-06-06",
                       harvest_date = "2014-10-01", # Guesstimated
                       initial_infection = "2014-06-26",# estimated "Early in the season" 20 days after sowing
                       seeding_rate = 40,
                       primary_infection_foci = expand.grid(x = 1:20,
                                                            y = 1:10,
                                                            load = 5),
                       spores_per_gp_per_wet_hour = sr
                     ) 
                   }
stopCluster(cl)

T2 <- Sys.time()
T2 - T1

```

```{r}
for(i in 1:length(perm_sr)) {
  i_1 <- summarise_trace(perm_sr[[i]])
  plot(
    y = i_1$infectious_gp / i_1$susceptible_gp,
    x = i_1$i_date,
    type = "l",
    ylim = c(0, 0.5)
  )
     if(i>1){
  i_2 <- summarise_trace(perm_sr[[i-1]])
  lines(y = i_2$infectious_gp/i_2$susceptible_gp,
       x = i_2$i_date,
       type = "l", col = "red")}
    if(i>2){
  i_3 <- summarise_trace(perm_sr[[i-2]])
  lines(y = i_3$infectious_gp/i_3$susceptible_gp,
       x = i_3$i_date,
       type = "l", col = "maroon")}
  
}
```

## Permutate fungicide sprays  

First I need to establish the main days when spores were spread.

`ta3` should show this in the dates when `exposed_gps` increased

```{r}
# find the change in exposed gp between days
s_ta3[,exposed_diff := exposed_gp - shift(exposed_gp)]

crutial_days <- s_ta3[exposed_diff >0, i_date]
```
now to create a function to randomly sample dates.
The function will not sample dates within a specified timeframe of each other and 
has the option of how many dates to sample.

```{r}
sample_dates <- function(dates,
                         n = 3,
                         min_diff = 5,
                         max_itter = 100) {
  for (i in 1:n) {
    itter <- 1
    if (i == 1) {
      out <- as.Date(vector(mode = "character", length = n))
      out[i] <- sample(dates,
                       size = 1)
      next
    }
    while (is.na(out[i])) {
      next_date <- sample(dates, size = 1)
      diff1 <- min(abs(next_date - out), na.rm = TRUE)
      if (diff1 >= min_diff) {
        out[i] <- next_date
      }
      itter <- itter + 1
        if(itter>= max_itter) break
    }
  }
  return(out)
}
f_dates <- sample_dates(crutial_days)
f_dates
```

Create list of fungicide applications  

```{r}
f_scenarios <-lapply(1:10, FUN = function(x) sample_dates(dates = crutial_days))
```



```{r}
for(q in 1:20){
  f_scenarios <-lapply(1:30, FUN = function(x) sample_dates(dates = crutial_days))

  T1 <- Sys.time()
  cl <- makeCluster(detectCores()-1)
  registerDoParallel(cl)

  foreach(
    fcide = f_scenarios,
    w = f_dat,
    .packages = "ascotraceR", 
    .inorder=FALSE) %dopar% {
      options(digits.secs=2)
      traced <-
        trace_asco(
          weather = f_dat,
          paddock_length = 10,
          paddock_width = 20,
          sowing_date = "2014-06-06",
          harvest_date = "2014-10-01",
          # Guesstimated
          initial_infection = "2014-06-26",
          # estimated "Early in the season" 20 days after sowing
          seeding_rate = 40,
          primary_infection_foci = expand.grid(x = 1:20,
                                               y = 1:10,
                                               load = 5),
          spores_per_gp_per_wet_hour = 0.6,
          fungicide_dates = fcide
        )
      l_dat <-
        list(f_dates = fcide,
             trace_summary = summarise_trace(traced))
      
      save("l_dat",
           file = paste0(
             "./Fungicide4_runs/",
             format(Sys.time(),
                    format = "%y%m%d_%H%M%OS"),
             "_perm_fg.Rdata"
           ))
    }
  
  stopCluster(cl)
  T2 <- Sys.time()
  T2 - T1
}

```

Sort by AUPDC

```{r}
aud1 <- unlist(lapply(perm_fg,FUN = function(x) summarise_trace(x)$AUDPC[1]))
perm_fg <- perm_fg[order(aud1)]
```



```{r}
for(i in 1:length(perm_fg)) {
  i_1 <- summarise_trace(perm_fg[[i]])
  plot(
    y = i_1$infectious_gp / i_1$susceptible_gp,
    x = i_1$i_date,
    type = "l",
    ylim = c(0, 0.5),
    xlab = "Date",
    ylab = "Percent infected growing points"
  )
  abline(v = f_scenarios[[i]], 
         col = "blue")
  text(x = as.Date("2014-06-30"),
       y = 0.45,
       labels = paste("AUDPC = ",unique(i_1$AUDPC)))
  #   if(i>1){
  # lines(y = perm_sr[[i-1]]$infectious_gp/perm_sr[[i-1]]$susceptible_gp,
  #      x = perm_sr[[i-1]]$i_date,
  #      type = "l", col = "red")}
  #   if(i>2){
  # lines(y = perm_sr[[i-1]]$infectious_gp/perm_sr[[i-1]]$susceptible_gp,
  #      x = perm_sr[[i-1]]$i_date,
  #      type = "l", col = "maroon")}
}
```


### Four spray scenarios

```{r}
for(q in 1:5) {
  # run preliminary simulation to find crutial days
  prelim <- trace_asco(
    weather = f_dat,
    paddock_length = 10,
    paddock_width = 20,
    sowing_date = "2014-06-06",
    harvest_date = "2014-10-30",
    # Guesstimated
    initial_infection = "2014-06-26",
    # estimated "Early in the season" 20 days after sowing
    seeding_rate = 50,
    # Desi
    primary_infection_foci = expand.grid(
      x = 1:20,
      y = 1:10,
      load = 5
    ),
    spores_per_gp_per_wet_hour = 0.6
  ) # kyabra = 0.6
  
  s_prelim <- summarise_trace(prelim)
  
  # find the change in exposed gp between days
  s_prelim[, exposed_diff := exposed_gp - shift(exposed_gp)]
  
  crutial_days <- s_prelim[exposed_diff > 0, i_date]
  
  f_scenarios <-lapply(1:100, FUN = function(x) sample_dates(dates = crutial_days,
                                                             n = 4))

  cl <- makeCluster(detectCores()-1)
  registerDoParallel(cl)

  foreach(
    fcide = f_scenarios,
    w = f_dat,
    .packages = "ascotraceR", 
    .inorder=FALSE) %dopar% {
      options(digits.secs=2)
      traced <-
        trace_asco(
          weather = f_dat,
          paddock_length = 10,
          paddock_width = 20,
          sowing_date = "2014-06-06",
          harvest_date = "2014-10-01",
          # Guesstimated
          initial_infection = "2014-06-26",
          # estimated "Early in the season" 20 days after sowing
          seeding_rate = 40,
          primary_infection_foci = expand.grid(x = 1:20,
                                               y = 1:10,
                                               load = 5),
          spores_per_gp_per_wet_hour = 0.6,
          fungicide_dates = fcide
        )
      l_dat <-
        list(f_dates = fcide,
             trace_summary = summarise_trace(traced))
      
      save("l_dat",
           file = paste0(
             "./Fungicide4_runs/",
             format(Sys.time(),
                    format = "%y%m%d_%H%M%OS"),
             "_perm_fg.Rdata"
           ))
    }
  
  stopCluster(cl)
}

}
```

